---
title: "Endangered species abundance"
author: "Mauro Zucconi"
date: "27/8/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary

The aim of this document is to explore the representation of taxa categoryes with at least one endangered species in some of the United States National Parks; first, in general, graphics include non vascular plants, invertebrates and other non-chordate taxa; second, I show the representation of the chordate taxa classified as endangered. Analysis were run exclusively for endangered species in some of the biggest Natural Parks in USA. Due to the lack of knowledge about encoding and the subsequent impossibility for this program to recognize the tilde, this document will be presented totally in english.

```{r, Data, echo=F, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(Rmisc) 
library(psych)
library(plyr)
library(car)
library(dplyr)
library(extrafont)
library(extrafontdb)
library(dismo)
library(knitr)
library(wesanderson)
######## Base de datos especies, cambio de nombre de columnas ########
parks <-read.csv("~/Desktop/Diversidad/Diversidad/parks.csv")
colnames(parks) <- c("ParkCode","ParkName", "State", "Acres", "Latitude", "Longitude")
species2 <-read_csv("~/Desktop/Diversidad/Diversidad/species.csv")
colnames(species2) <- c("SpeciesID","ParkName", "Category", "Order", "Family", "SName", 
                        "CommonNames", "RecordStatus", "Occurrence", "Nativeness", "Abundance", 
                        "Seasonality", "ConservationStatus", "X14")
```
```{r, echo=F, warning=F, message=FALSE}
aa <- species2 %>% count(c("ParkName", "SName", "Category", "ConservationStatus")) %>% 
  filter(!is.na("Abundance")) %>% filter(!is.na("Occurrence")) %>%  filter(!is.na("ConservationStatus")) %>% 
  filter(ConservationStatus == "Endangered") %>% filter(Category == "Mammal" | Category == "Bird"| Category == "Reptile" |Category == "Amphibian" | Category == "Fish"| Category == "Insect" | Category == "Invertebrate")
aaa <- species2 %>% count(c("ParkName", "SName", "Category", "ConservationStatus")) %>% 
  filter(!is.na("Abundance")) %>% filter(!is.na("Occurrence")) %>% filter(!is.na("ConservationStatus"))
aa1<- aa %>% group_by(ParkName, Category, ConservationStatus) %>% dplyr::summarise(n = n_distinct(SName))
aaa1 <-aaa %>% group_by(ParkName, Category, ConservationStatus) %>% dplyr::summarise(n = n_distinct(SName)) %>% filter(!is.na("ConservationStatus"))
tot <- full_join(parks,aa1,by="ParkName",copy=FALSE)
tot1 <-tot %>% filter(!is.na(ConservationStatus)) %>% filter(!is.na(Category))
tat <- full_join(parks,aaa1,by="ParkName",copy=FALSE)
tat1 <-tat %>% filter(!is.na(ConservationStatus)) %>% filter(!is.na(Category))
wer <- tat1 %>% group_by(ParkName) %>% summarise_all(funs(mean, sd)) %>% filter(!is.na(n_sd)) %>% dplyr::select(ParkName, n_mean, n_sd)
colnames(wer) <- c("Park Name", "Mean abundance", "Standard-Deviation abundance")

```
```{r, echo=F, warning=F, message=F, cap=c("Mean number of categoryes classified as endangered in each Natural Park")}
totabla<-tot1 %>% group_by(Category, ParkName) %>%  dplyr::summarise(mean = mean(n),n = n())
kable(totabla, caption = "Table 1. Mean number of categoryes with species classified as endangered in each Natural Park")

```


```{r, echo=F, warning=F, message=F, fig.width = 15, fig.asp = .62, fig.cap=c("Fig.1: Taxa categories with at least one species categorized as endangered in each Park", "Fig.2: Chordata taxa categories with at least one species categorized as endangered in each Park", "Fig.3:Graphic relationship between the latitude and the longitud with the chordata taxa", "Fig.4: Graphic relationship between the latitude and the longitud with the area of the park (in acres)") }


ggplot(tat1,aes(x = ParkName, y = n)) + 
  geom_col(aes(fill = Category))+ 
  scale_fill_manual(values = c(rep(c("lightgreen", "lightgray", "steelblue", "black", "skyblue","#E69F00","darkgreen", "red", "lightgray", "darkgreen", "khaki", "pink", "yellowgreen"))))+
  theme_bw()+
  theme(panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        strip.background=element_blank())+
  ylab("Frequency")+ xlab("Park Name")+
  theme(text=element_text(family="Times New Roman"), 
        axis.line=element_line(colour = "black"),
        axis.text.x = element_text(angle=60, hjust=1),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(), 
        panel.border=element_blank(),
        panel.background=element_blank(), 
        legend.title=element_text(size=18, face= "bold",colour="black"), 
        legend.text=element_text(size=15),
        axis.title=element_text(size=18, face= "bold",colour="black"),
        axis.text=element_text(size=15))+
  scale_x_discrete(labels = c("Acadia NP","Arches NP","Badlands NP", "Big Bend NP","Biscayne NP", "Black Canyon NP", "Bryce Canyon NP","Canyonlands NP", "Capitol Reef NP",  "Carlsbad Caverns NP","Channel Islands NP", "Congaree NP", "Crater Lake NP", "Cuyahoga Valley NP", "Death Valley NP","Dry Tortugas NP ", "Everglades NP", "Glacier Bay NP","Grand Canyon NP", "Grand Teton NP", "Great Basin NP", "Great Smoky Mountains NP", "Guadalupe Mountains NP", 
"Haleakala NP", "Hawaii Volcanoes NP", "Hot Springs NP", "Isle Royale NP", "Joshua Tree NP", "Katmai NP", "Kenai Fjords NP", "Lake Clark NP", "Lassen Volcanic NP", "Mammoth Cave NP", "Mesa Verde NP", "Mount Rainier NP",  "North Cascades NP", "Olympic NP", "Petrified Forest NP", "Pinnacles NP","Redwood NP", "Rocky Mountain NP", "Saguaro NP", "Sequoia-Kings Canyon NPs", "Shenandoah NP","Theodore Roosevelt NP", "Voyageurs NP", "Wind Cave NP", "Wrangell-St Elias NP", "Yellowstone NP", "Yosemite NP", "Zion NP"))




ggplot(tot1,aes(x = ParkName, y = n)) + 
  geom_col(aes(fill = Category))+
  scale_fill_manual(values = c(rep(c("lightgreen", "lightgray", "steelblue", "black", "skyblue","#E69F00","darkgreen"))))+
  theme_bw()+
  theme(panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        strip.background=element_blank())+
  ylab("Frequency")+ xlab("")+
  theme(text=element_text(family="Times New Roman"), 
        axis.line=element_line(colour = "black"),
        axis.text.x = element_text(angle=60, hjust=1),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(), 
        panel.border=element_blank(),
        panel.background=element_blank(), 
        legend.title=element_text(size=18, face= "bold",colour="black"), 
        legend.text=element_text(size=15),
        axis.title=element_text(size=18, face= "bold",colour="black"),
        axis.text=element_text(size=15))+
        scale_x_discrete(labels = c("Acadia NP","Arches NP","Badlands NP", "Big Bend NP","Biscayne NP", "Black Canyon NP", "Bryce Canyon NP","Canyonlands NP", "Capitol Reef NP",  "Carlsbad Caverns NP","Channel Islands NP", "Congaree NP", "Crater Lake NP", "Cuyahoga Valley NP", "Death Valley NP","Dry Tortugas NP ", "Everglades NP", "Glacier Bay NP","Grand Canyon NP", "Grand Teton NP", "Great Basin NP", "Great Smoky Mountains NP", "Guadalupe Mountains NP", 
"Haleakala NP", "Hawaii Volcanoes NP", "Hot Springs NP", "Isle Royale NP", "Joshua Tree NP", "Katmai NP", "Kenai Fjords NP", "Lake Clark NP", "Lassen Volcanic NP", "Mammoth Cave NP", "Mesa Verde NP", "Mount Rainier NP",  "North Cascades NP", "Olympic NP", "Petrified Forest NP", "Pinnacles NP","Redwood NP", "Rocky Mountain NP", "Saguaro NP", "Sequoia-Kings Canyon NPs", "Shenandoah NP","Theodore Roosevelt NP", "Voyageurs NP", "Wind Cave NP", "Wrangell-St Elias NP", "Yellowstone NP", "Yosemite NP", "Zion NP"))




ggplot(tot1,aes(x = Longitude, y = Latitude)) + 
geom_point(aes(color = Category), size = 5)+
scale_color_manual(values = c(rep(c("lightgreen", "lightgray", "steelblue", "black", "skyblue","#E69F00","darkgreen"))))+
  theme_bw()+
  theme(panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        strip.background=element_blank())+
  ylab("Latitud")+ xlab("Longitud")+
  theme(text=element_text(family="Times New Roman"), 
        axis.line=element_line(colour = "black"),
        axis.text.x = element_text(angle=0, hjust=1),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(), 
        panel.border=element_blank(),
        panel.background=element_blank(), 
        legend.title=element_text(size=18, face= "bold",colour="black"), 
        legend.text=element_text(size=15),
        axis.title=element_text(size=18, face= "bold",colour="black"),
        axis.text=element_text(size=15))




ggplot(tot1,aes(x = Longitude, y = Latitude))+ 
  geom_point(aes(size = Acres))+
  scale_color_manual(values=wes_palette(n=4, name="GrandBudapest1")) +
  theme_bw()+
  theme(panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        strip.background=element_blank())+
  ylab("Latitud")+ xlab("Longitud")+
  theme(text=element_text(family="Times New Roman"), 
        axis.line=element_line(colour = "black"),
        axis.text.x = element_text(angle=0, hjust=1),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(), 
        panel.border=element_blank(),
        panel.background=element_blank(), 
        legend.title=element_text(size=18, face= "bold",colour="black"), 
        legend.text=element_text(size=15),
        axis.title=element_text(size=18, face= "bold",colour="black"),
        axis.text=element_text(size=15))


